Toward Supervised Reinforcement Learning with Partial States for Social HRI

### 2017-08-22 paper.tex
to fu

\maketitle \begin{abstract}
%As robots will integrate human society, they will have to satify humans'
    %expectations. However at the design time, it seems hardly possible to know
    %in advance exactly all the behaviours expected from the robot. As such to
    %be accepted in human society robots have to learn how to interact with
    %humans.
    Social interacting is a complex task for which machine learning holds
    particular promise. However, as no sufficiently accurate simulator of human
    interactions exists today, the learning of social interaction strategies has
    to happen online in the real world. Actions executed by the robot impact on
    humans, and as such have to be carefully selected, making it impossible to
    rely on random exploration. Additionally, no clear reward function exists
    for social interactions. This implies that traditional approaches used for
    Reinforcement Learning cannot be directly applied for learning how to
    interact with the social world. As such we argue that robots will profit
    from human expertise and guidance to learn social interactions. And as the
    quantity of input a human can provide is limited, new methods have to be
    designed to use human input more efficiently. In this paper we describe a
    setup in which we combine a framework called Supervised Progressively
    Autonomous Robot Competencies (SPARC), which allows safer online learning
    with Reinforcement Learning, with the use partial states rather than full
    states to accelerate generalisation and obtain more quickly a usable action
    policy.  
\end{abstract}
%===============================================================================

@@

Human-Robot Interaction (HRI) studies how people and robot can co-exist in
 society, and how they can interact socially in different environments and
 contexts. Robot are expected to behave appropriately regardless of the domain
 of interaction. However, it is impossible to foresee all possible interaction
 outcomes in dynamic and open social domains, as such the robot's responses
 cannot be implemented in the robot before its deployment in the real world.
 Similarly to  people, robots need to be able to learn how to complete tasks
 throug creating and optimising action policies. This includes learning social
 norms and how to make sense of the social world.

@@

%These relevant features in the state space can be encoded in the representation of
%the space the robot uses for action selection or be learnt. Recent successes in Deep
%Learning~\cite{lecun2015deep} have shown that complex features can be extracted
%by neural networks assuming that enough datapoints are available.

@@

===============================================================================

@@

%Classic approaches in RL generally rely on exploring virtual environments where
%the agent tries to maximise the reward.
In many cases where RL achieves success, the agent has access to a virtual
environment where the only real cost of exploring is computational effort: the
agent can interact as long as needed to gather enough information on the
environment and the result of its actions on the environment in order to find a
sufficiently optimal action policy. Using simulation RL can achieve impressive
results, as shown in RL learning to play Backgammon in the
90s~\cite{tesauro1995temporal} to the more recent success in mastering the game
of Go~\cite{silver2016mastering}. Even when a virtual environment is available,
but especially when it is not, human knowledge and effort is required during the
design of the algorithm or the learning phase before the learning can be
successful.

@@

\subsubsection{Design decisions:}
Initial knowledge has to be used in the design of the algorithm, the
representation of the state and actions spaces and the reward funtion. For
example, only carefully crafted features allowed
Tesauro to improve its algorithm for Backgammon from a intermediate-level player
to super-human level~\cite{tesauro1995temporal}. Similarly, the design of
complex neural networks for state generalisation, the representation of actions
as motor primitives rather than raw motor angles or adding additional
information in the reward function have important impacts on the ability of
the robot to reach a successful policy~\cite{kober2013reinforcement}.

\subsubsection{Demonstrations:} 
Initial knowledge can be provided to the agent
through demonstrations. These demonstrations can be used to create an initial
policy which is sufficiently efficient to start interacting in the environment
and gather information to improve over time. This initial policy, required to
receive meaningful feedback from the environment can be impossible to reach by
relying only on random exploration and feedback from the environment. For
example the game of Go in its larger board size contains more than $10^{210}$
states, so exhaustive search is not possible.  Silver et al.
\cite{silver2016mastering} started with supervised learning from Go masters'
games to learn a decent enough policy and then proceeded using deep learning,
self play and tree search to achieve super-human capabilities and beating the
best human players.

@@

\subsubsection{Guiding the learning:}
Rather than solely providing initial knowledge to the agents, humans can also
guide the agent during its learning, a method more resembling human teaching.

A first approach consists in sequencing the tasks the robot will face.  This
approach, known as ``scaffolding'' \cite{saunders2006teaching}, progressively
increases the difficulty and complexity of the task as the robot is learning to
reach policies which would take prohibitively long without scaffolding.

@@

In \cite{senft2015sparc}, we introduced the Supervised Progressively Autonomous
Robot Competencies (SPARC). SPARC relies on a supervisor with the ability to
control the robot's actions. This supervisor is presented with the action the
robot is about to execute. He can then choose to cancel it, allow it to be
executed, or can select an alternative action. Based on the supervisor's
decisions, the robot learns which actions are desired and can as such refine its
policy over time reducing the need for the supervisor to correct or select an
action.

@@

%SPARC has been designed with the idea of teaching a robot to interact with humans,
%it is specially usefull when the frequency of action selection is low (less than
%1Hz) so action can be evaluated by a human before being executed, when the agent
%has to learn with a limited number of datapoints and when
%undesired actions executed by an agent can have a high cost and are to be
%avoided.

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot, making the
assumption that every action has been explicitly or implicitely approved by the
supervisor. However this method was  directly mapping a single (state, action)
pair to a reward without making use of any kind of generalisation. As such it
was not applicable to environments with a continuous or high dimensional space
or non-deterministic transitions from one state to another, elements which are
typically present in social interactions.

@@

%In \cite{sutton1996generalisation}, Sutton proposed to use sparse coarse
%coding
%dividing the state in slices and tiles as a way to generalise quickly for
%continuous spaces in RL.

@@

SPARC has been shown to be compatible with RL in \cite{senft2017supervised}, and
can also be easily combined with the approach presented in this paper using
partial states. For example, when selecting an action for the robot to be
executed, the supervisor can also select which features in the environment
should be selected as the partial state, and this associates the reward to this
action in this partial state. Similarly, when an action is proposed to the
supervisor the features represented by the dimensions of the closest partial
state for this action can be exposed to the supervisor as a way to explain why
this action has been selected. Facing this, the supervisor can: (1) not react,
allowing the action to be executed and associating a reward of +1 to the partial
state and the action proposed, (2) change the partial state to correct the
features related to this action or (3) cancel it, preventing the execution and
associating a reward of -1 to the partial state identified by the robot or the
supervisor.

@@

The GUI (Figure \ref{fig:gui}) is an augmented
version of the game itself, which can be use to make the robot move items on the
game by dragging them on the GUI or which can display actions proposed by the
robot with a cancel button to refuse an action. For example in
Figure~\ref{fig:gui} the robot proposes to move the eagle to the rat, and
highlight (as shown by blue circles) the eagle and the rat. This indicate that
features relevant to the eagle and the rat have been used to select this action.

@@

The current work could also be extended to allow the agent to continue to
    improve its behaviour even in the absence of a supervisor, progressively
    exploring around the learnt policy improving its behaviour beyond the
    demonstrations. This could be done by allowing the supervisor to provide
    rewards during the learning phase and combine these rewards with the
    demonstrations to learn a reward function in a fashion similar to Inverse
    Reinforcement Learning \cite{abbeel2004apprenticeship} or TAMER
    \cite{knox2009interactively}. This could also use partial states associated
    to these rewards to ease the generalisation of the reward function with only
    a low number of datapoints.  But without the supervisor, the assumption that
    only a myopic action selection is sufficient would not hold anymore and the
    problem of delayed rewards would have to be tackled.
%===============================================================================
    %\section{Conclusion} \label{sec:conclusion}
    %==============================================================================
\section{Acknowledgments} This work was supported by the EU FP7 DREAM project
    (grant no.  611391) and EU H2020 Marie Sklodowska-Curie Actions project
    DoRoThy (grant 657227).

>>>>

\maketitle

\begin{abstract}

    Social interacting is a complex task for which machine learning holds
    particular promise. However, as no sufficiently accurate simulator of human
    interactions exists today, the learning of social interaction strategies has
    to happen online in the real world. Actions executed by the robot impact on
    humans, and as such have to be carefully selected, making it impossible to
    rely on random exploration. Additionally, no clear reward function exists
    for social interactions. This implies that traditional approaches used for
    Reinforcement Learning cannot be directly applied for learning how to
    interact with the social world. As such we argue that robots will profit
    from human expertise and guidance to learn social interactions. And as the
    quantity of input a human can provide is limited, new methods have to be
    designed to use human input more efficiently. In this paper we describe a
    setup in which we combine a framework called Supervised Progressively
    Autonomous Robot Competencies (SPARC), which allows safer online learning
    with Reinforcement Learning, with the use partial states rather than full
    states to accelerate generalisation and obtain more quickly a usable action
    policy.

\end{abstract}
%===============================================================================

@@

Human-Robot Interaction (HRI) studies how people and robot can co-exist in
 society, and how they can interact socially in different environments and
 contexts. Robot are expected to behave appropriately regardless of the domain
 of interaction. However, it is impossible to foresee all possible interaction
 outcomes in dynamic and open social domains, as such the robot's responses
 cannot be implemented in the robot before its deployment in the real world.
 Similarly to  people, robots need to be able to learn how to complete tasks
 through creating and optimising action policies. This includes learning social
 norms and how to make sense of the social world.

@@

@@

%===============================================================================

@@

In many cases where RL achieves success, the agent has access to a virtual
environment where the only real cost of exploring is computational effort: the
agent can interact as long as needed to gather enough information on the
environment and the result of its actions on the environment in order to find a
sufficiently optimal action policy. Using simulation RL can achieve impressive
results, as shown in RL learning to play Backgammon in the
90s~\cite{tesauro1995temporal} to the more recent success in mastering the game
of Go~\cite{silver2016mastering}. Even when a virtual environment is available,
but especially when it is not, human knowledge and effort is required during the
design of the algorithm or the learning phase before the learning can be
successful.

@@

\subsubsection{Design decisions:}

Initial knowledge has to be used in the design of the algorithm, the
representation of the state and actions spaces and the reward funtion. For
example, only carefully crafted features allowed Tesauro to improve its
algorithm for Backgammon from a intermediate-level player to super-human
level~\cite{tesauro1995temporal}. Similarly, the design of complex neural
networks for state generalisation, the representation of actions as motor
primitives rather than raw motor angles or adding additional information in the
reward function have important impacts on the ability of the robot to reach a
successful policy~\cite{kober2013reinforcement}.

\subsubsection{Demonstrations:}

Initial knowledge can be provided to the agent through demonstrations. These
demonstrations can be used to create an initial policy which is sufficiently
efficient to start interacting in the environment and gather information to
improve over time. This initial policy, required to receive meaningful feedback
from the environment can be impossible to reach by relying only on random
exploration and feedback from the environment. For example the game of Go in its
larger board size contains more than $10^{210}$ states, so exhaustive search is
not possible.  Silver et al.~\cite{silver2016mastering} started with supervised
learning from Go masters' games to learn a decent enough policy and then
proceeded using deep learning, self play and tree search to achieve super-human
capabilities and beating the best human players.

@@

\subsubsection{Guiding the learning:}

Rather than solely providing initial knowledge to the agents, humans can also
guide the agent during its learning, a method more resembling human teaching.

A first approach consists in sequencing the tasks the robot will face.  This
approach, known as ``scaffolding''~\cite{saunders2006teaching}, progressively
increases the difficulty and complexity of the task as the robot is learning to
reach policies which would take prohibitively long without scaffolding.

@@

In~\cite{senft2015sparc}, we introduced the Supervised Progressively Autonomous
Robot Competencies (SPARC). SPARC relies on a supervisor with the ability to
control the robot's actions. This supervisor is presented with the action the
robot is about to execute. He can then choose to cancel it, allow it to be
executed, or can select an alternative action. Based on the supervisor's
decisions, the robot learns which actions are desired and can as such refine its
policy over time reducing the need for the supervisor to correct or select an
action.

@@

In~\cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot, making the
assumption that every action has been explicitly or implicitely approved by the
supervisor. However this method was directly mapping a single (state, action)
pair to a reward without making use of any kind of generalisation. As such it
was not applicable to environments with a continuous or high dimensional space
or non-deterministic transitions from one state to another, elements which are
typically present in social interactions.

@@

@@

SPARC has been shown to be compatible with RL in~\cite{senft2017supervised}, and
can also be easily combined with the approach presented in this paper using
partial states. For example, when selecting an action for the robot to be
executed, the supervisor can also select which features in the environment
should be selected as the partial state, and this associates the reward to this
action in this partial state. Similarly, when an action is proposed to the
supervisor the features represented by the dimensions of the closest partial
state for this action can be exposed to the supervisor as a way to explain why
this action has been selected. Facing this, the supervisor can: (1) not react,
allowing the action to be executed and associating a reward of +1 to the partial
state and the action proposed, (2) change the partial state to correct the
features related to this action or (3) cancel it, preventing the execution and
associating a reward of -1 to the partial state identified by the robot or the
supervisor.

@@

The GUI (Figure~\ref{fig:gui}) is an augmented version of the game itself, which
can be use to make the robot move items on the game by dragging them on the GUI
or which can display actions proposed by the robot with a cancel button to
refuse an action. For example in Figure~\ref{fig:gui} the robot proposes to move
the eagle to the rat, and highlight (as shown by blue circles) the eagle and the
rat. This indicate that features relevant to the eagle and the rat have been
used to select this action.

@@

The current work could also be extended to allow the agent to continue to
improve its behaviour even in the absence of a supervisor, progressively
exploring around the learnt policy improving its behaviour beyond the
demonstrations. This could be done by allowing the supervisor to provide rewards
during the learning phase and combine these rewards with the demonstrations to
learn a reward function in a fashion similar to Inverse Reinforcement
Learning~\cite{abbeel2004apprenticeship} or TAMER ~\cite{knox2009interactively}.
This could also use partial states associated to these rewards to ease the
generalisation of the reward function with only a low number of datapoints.  But
without the supervisor, the assumption that only a myopic action selection is
sufficient would not hold anymore and the problem of delayed rewards would have
to be tackled.

%==============================================================================
\section{Acknowledgments} This work was supported by the EU FP7 DREAM project
(grant no.  611391) and EU H2020 Marie Sklodowska-Curie Actions project DoRoThy
(grant 657227).
